---
title: "Module 4 Assignment 2"
author: "Lynsey Scheneman"
date: '2022-09-20'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lpSolveAPI)
```

###
The Weigelt Corporation has three branch plants with excess production capacity. Fortunately, the corporation has a new product ready to begin production, and all three plants have this capability, so some of the excess capacity can be used in this way. This product can be made in three sizes--large, medium, and small--that yield a net unit profit of $420, $360, and $300, respectively. Plants 1, 2, and 3 have the excess capacity to produce 750, 900, and 450 units per day of this product, respectively, regardless of the size or combination of sizes involved. The amount of available in-process storage space also imposes a limitation on the production rates of the new product. Plants 1, 2, and 3 have 13,000, 12,000, and 5,000 square feet, respectively, of in-process storage space available for a day's production of this product. Each unit of the large, medium, and small sizes produced per day requires 20, 15, and 12 square feet, respectively. Sales forecasts indicate that if available, 900, 1,200, and 750 units of the large, medium, and small sizes, respectively, would be sold per day. At each plant, some employees will need to be laid off unless most of the plant’s excess production capacity can be used to produce the new product. To avoid layoffs if possible, management has decided that the plants should use the same percentage of their excess capacity to produce the new product.Management wishes to know how much of each of the sizes should be produced by each of the plants to maximize profit.

# Define the decision variables 
The decision variables are the # of units of each new product (small, medium and large) that should be produced on each plant (1, 2, and 3) to maximize the company’s profit.

x1 = small product, plant 1
x2 = medium product, plant 1
x3 = large product, plant 1
x4 = small product, plant 2
x5 = medium product, plant 2
x6 = large product, plant 2
x7 = small product, plant 3
x8 = medium product, plant 3
x9 = large product, plant 3

# Formulate a linear programming model for this problem.

Let 
x1 = small product, plant 1
x2 = medium product, plant 1
x3 = large product, plant 1
x4 = small product, plant 2
x5 = medium product, plant 2
x6 = large product, plant 2
x7 = small product, plant 3
x8 = medium product, plant 3
x9 = large product, plant 3

Max
Z = 300(x1 + x4 + x7) + 360(x2 + x5 + x8) + 420(x3 + x6 + x9)

Subject to 

Excess Capacity 
x1 + x2 + x3 <= 750 
x4 + x5 + x6 <= 900
x7 + x8 + x9 <= 450

In-process Storage Space (SQ FT)
12(x1) + 15(x2) + 20(x3) <= 13000
12(x4) + 15(x5) + 20(x6) <= 12000
12(x7) + 15(x8) + 20(x9) <= 5000

Sales Forecast 
x1 + x4 + x7 <= 750
x2 + x5 + x8 <= 1200
x3 + x6 + x9 <= 900

Percentage 
900(x1 + x2 + x3) - 750(x4 + x5 + x6) = 0
450(x4 + x5 + x6) - 900(x7 + x8 + x9) = 0
450(x1 + x2 + x3) - 750(x7 + x8 + x9) = 0

Other
x1,x2, x3, x4, x5, x6, x7, x8, x9 >= 0

#Solve, identify variables and objective funtion
```{r}
dvar <- make.lp(0,9)
set.objfn(dvar,c(420,420,420,360,360,360,300,300,300))
lp.control(dvar,sense='max')
```
#Add constraints
Excess Capacity
```{r}
add.constraint(dvar, c(1,0,0,1,0,0,1,0,0), "<=", 750)
add.constraint(dvar, c(0,1,0,0,1,0,0,1,0), "<=", 900)
add.constraint(dvar, c(0,0,1,0,0,1,0,0,1), "<=", 450)
```

#
Storage Space (SQ FT)
```{r}
add.constraint(dvar, c(20,0,0,15,0,0,12,0,0), "<=", 13000)
add.constraint(dvar, c(0,20,0,0,15,0,0,12,0), "<=", 12000)
add.constraint(dvar, c(0,0,20,0,0,15,0,0,12), "<=", 5000)
```

#
Sales Forecast
```{r}
add.constraint(dvar, c(1,1,1,0,0,0,0,0,0), "<=", 900)
add.constraint(dvar, c(0,0,0,1,1,1,0,0,0), "<=", 1200)
add.constraint(dvar, c(0,0,0,0,0,0,1,1,1), "<=", 750)
```

#
Percentage
```{r}
add.constraint(dvar, c(0.001333,-0.00111,0,0.001333,-0.00111,0,0.001333,-0.00111,0), "=", 0)
add.constraint(dvar, c(0,-0.00111,0.00222,0,-0.00111,0.00222,0,-0.00111,0.00222), "=", 0)
```

#
Set up linear Program and ensure program ran correctly
```{r}
Rows <- c("1", "2", "3", "4", "5,", "6", "7", "8", "9", "10", "11")
Columns <- c("1", "2", "3", "4", "5,", "6", "7", "8", "9")
dimnames(dvar) <- list(Rows,Columns)
dvar
solve(dvar)
```

#
Get answers
```{r}
get.objective(dvar)
```

```{r}
get.variables(dvar)
```

```{r}
get.constraints(dvar)
```

#Description of Results
Maximum Profit = $695,906.2
To achieve this maximum profit, the company needs to produce:
- Large: Plant 1 at 518 units, Plant 2 at 0 units, Plant 3 at 0 units
- Medium: Plant 1 at 176 units, Plant 2 at 667 units, Plant 3 at 0 units
- Small: Plant 1 at 0 units, Plant 2 at 167 units, Plant 3 at 417 units

**ALL QUANTITIES HAVE BEEN ROUNDED UP BECAUSE THEY CAN ONLY PRODUCE WHOLE UNITS

