---
title: "Assignment 3, Module 6"
author: "Lynsey Scheneman"
date: "2022-10-4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lpSolveAPI)
```

Heart Start produces automated external defibrillators (AEDs) in each of two different plants (A and B). The unit production costs and monthly production capacity of the two plants are indicated in the table below. The AEDs are sold through three wholesalers. The shipping cost from each plant to the warehouse of each wholesaler along with the monthly demand from each wholesaler are also indicated in the table. How many AEDs should be produced in each plant, and how should they be distributed to each of the three wholesaler warehouses so as to minimizethe combined cost of production and shipping?
Formulate and solve this transportation problem.

# Define the decision variables 
x1 = plant A, warehouse 1
x2 = plant A, warehouse 2
x3 = plant A, warehouse 3
x4 = plant A, dummy warehouse
x5 = plant B, warehouse 1
x6 = plant B, warehouse 2
x7 = plant B, warehouse 3
x8 = plant B, dummy warehouse

Supply is 220 and demand is 210. Because this is the case, we will use a dummy warehouse that has a monthly demand of 10. Thus, supply (220) = demand (220).

#Define the objective function

MIN
Y = 22(x1) + 14(x2) + 30(x3) + 16(x5) + 20(x6) + 24(x7) + 600(x1 + x2 + x3 + x4) + 625(x5 + x6 + x7 + x8)



#Define the constraints
Production Capacity
x1 + x2 + x3 + x4 = 100
x5 + x6 + x7+ x8 = 120

Monthly Demand
x1 + x5 = 80
x2 + x6 = 60 
x3 + x7 = 70 
x4 + x8 = 10 

#Solve, identify variables and objective funtion

```{r}
dvar <- make.lp(0,8)
set.objfn(dvar,c(622,614,630,0,641,645,649,0))
lp.control(dvar,sense='min')
```

#Add constraints
Production Capacity 
```{r}
add.constraint(dvar, c(1, 1, 1, 1, 0, 0, 0, 0), "=", 100)
add.constraint(dvar, c(0, 0, 0, 0, 1, 1, 1, 1), "=", 120)
```

Monthly Demand
```{r}
add.constraint(dvar, c(1, 0, 0, 0, 1, 0, 0, 0), "=", 80)
add.constraint(dvar, c(0, 1, 0, 0, 0, 1, 0, 0), "=", 60)
add.constraint(dvar, c(0, 0, 1, 0, 0, 0, 1, 0), "=", 70)
add.constraint(dvar, c(0, 0, 0, 1, 0, 0, 0, 1), "=", 10)
```

#Solve and Describe Results
```{r}
solve(dvar)
```

```{r}
get.objective(dvar)
get.variables(dvar)
get.constraints(dvar)
```

Results: 
Costs will equal $132,790

x1 = 0 units
x2 = 60 units
x3 = 40 units
x4 = 0 units
x5 = 80 units
x6 = 0 units
x7 = 30 units
x8 = 10 units (dummy capacity)