---
title: "Assignment 4, Module 8"
author: "Lynsey Scheneman"
date: "2022-10-21"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

library(lpSolveAPI)
library(Benchmarking)

knitr::opts_chunk$set(echo = TRUE)
```

#Question 1: Formulate and perform DEA analysis under all DEA assumptions of FDH, CRS, VRS, IRS, DRS, and FRH
Create Table
```{r}
x <- matrix(c(150,400,320,520,350,320,0.2,0.7,1.2,2.0,1.2,0.7), ncol = 2)
y <- matrix(c(14000,14000,42000,28000,19000,14000, 3500,21000,10500,42000,25000,15000),ncol = 2)
colnames(x) <- c("Staff Hours per Day", "Supplies per Day")
colnames(y) <- c("Reimbursed Patient Days", "Privately Paid Patient Days")

x
y
```

Perform DEA analysis under the assumptions
```{r}
FDH <- dea(x, y, RTS = "fdh")
CRS <- dea(x, y, RTS = "crs")
VRS <- dea(x, y, RTS = "vrs")
IRS <- dea(x, y, RTS = "irs")
DRS <- dea(x, y, RTS = "drs")
FRH <- dea(x, y, RTS = "add")
```

#Question 2: Determine the Peers and Lambdas under each of the above assumptions
```{r}
FDH
peers(FDH)
lambda(FDH)
dea.plot.isoquant(x, y, RTS = "fdh")
```

```{r}
CRS
peers(CRS)
lambda(CRS)
dea.plot.isoquant(x,y,RTS="crs") 
```

```{r}
VRS
peers(VRS)
lambda(VRS)
dea.plot.isoquant(x,y,RTS="vrs")
```

```{r}
IRS
peers(IRS)
lambda(IRS)
dea.plot.isoquant(x,y,RTS="irs") 
```

```{r}
DRS
peers(DRS)
lambda(DRS)
dea.plot.isoquant(x,y,RTS="drs") 
```

```{r}
FRH
peers(FRH)
lambda(FRH)
```

#Question 3: Summarize your results in a tabular format
```{r}
Summary <- read.csv("~/Desktop/Summary.csv", header=FALSE)
Summary
```

#Question 4: Compare and contrast the above results
After looking at the 6 DEA models, it can be determined that: 
-under FDH and FRH, all facilities are efficent
-under CRS and DRS, facilities 1-4 are efficent but facility 5 runs ar 97.75% and facility 6 runs ar 86.75%
- under VRS and IRS, facilites 1-5 are efficent but facility 6 runs at 89.63 percent.

