---
title: "schenem_6"
author: "Lynsey Scheneman"
date: "2022-11-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lpSolveAPI)
```

AP is a shipping service that guarantees overnight delivery of packages in the continental US.The company has various hubs at major cities and airports across the country. Packages are received at hubs, and then shipped to intermediate hubs or to their final destination.The manager of the AP hub in Cleveland is concerned about labor costs, and is interested in determining the most effective way to schedule workers. The hub operates seven days a week, and the number of packages it handles varies from one day to another. The table below provides an estimate of the number of workers needed each day of the week.
Sunday = 18 Workers 
Monday = 27 Workers
Tuesday = 22 Workers 
Wednesday = 26 Workers 
Thursday = 25 Workers
Friday = 21 Workers
Saturday = 19 Workers
Package handlers at AP are guaranteed a five-day work week with two consecutive days off. The base wage for the handlers is $750 per week. Workers working on Saturday or Sunday receive an additional $25 per day. The possible shifts and salaries for package handlers are:
Shift 1, $775
Shift 2, $800
Shift 3, $800
Shift 4, $800
Shift 5, $800
Shift 6, $775
Shift 7, $750

Q: The manager wants to keep the total wage expenses as low as possible while ensuring that there are sufficient number of workers available each day. Formulate and solve the problem. What was the total cost? How many workers are available each day? Hint: The number of available workers each day can exceed, but can not be below the required
amount.

#Formulate Problem

x1 = Shift 1
x2 = Shift 2
x3 = Shift 3       
x4 = Shift 4
x5 = Shift 5
x6 = Shift 6
x7 = Shift 7

MIN: 775(x1) + 800(x2) + 800(x3) + 800(x4) + 800(x5) + 775(x6) + 750(x7)

Subject to:

    X2+ X3+ X4+ X5+ X6     >= 18  Sunday
    
        X3+ X4+ X5+ X6+ X7 >= 27  Monday
        
X1+         X4+ X5+ X6+ X7 >= 22  Tuesday

X1+ X2+         X5+ X6+ X7 >= 26  Wednesday
  
X1+ X2+ X3+         X6+ X7 >= 25  Thursday
  
X1+ X2+ X3+ X4+         X7 >= 21   Friday
  
X1+ X2+ X3+ X4+ X5         >= 19   Saturday
  
#Solve Problem
```{r}
dvar <- make.lp(7,7)
set.objfn(dvar, c(775, 800, 800, 800, 800, 775, 750))
lp.control(dvar, sense = 'min')
```

```{r}
add.constraint(dvar, c(0,1,1,1,1,1,0), ">=", 18)
add.constraint(dvar, c(0,0,1,1,1,1,1), ">=", 27)
add.constraint(dvar, c(1,0,0,1,1,1,1), ">=", 22)
add.constraint(dvar, c(1,1,0,0,1,1,1), ">=", 26)
add.constraint(dvar, c(1,1,1,0,0,1,1), ">=", 25)
add.constraint(dvar, c(1,1,1,1,0,0,1), ">=", 21)
add.constraint(dvar, c(1,1,1,1,1,0,0), ">=", 19)
```

```{r}
solve(dvar)
get.objective(dvar)
get.variables(dvar)
```

#Explaination of Results
Minimum Total Cost: $25,175 
Availiable Workers Each Day: (rounded up to account for a whole person)
x1 = Shift 1 = 2 Workers
x2 = Shift 2 = 4 Workers
x3 = Shift 3 = 7 Workers
x4 = Shift 4 = 0 Workers
x5 = Shift 5 = 8 Workers
x6 = Shift 6 = 1 Worker
x7 = Shift 7 = 13 Workers